---
title: "Introduction to lr"
author: "Youqi Yang"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to lr}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(linearmodel)
library(NHANES)
library(dplyr)
```

## Data: NHANES

To explore the applications of our function, we will use a subset of a R built-in data in package(NHANES) `NHANES`.

```{r}
data(NHANES)
data <- NHANES |>
  select("Age", "Race1", "Weight", "Height")

data <- na.omit(data)
head(data)
dim(data)
```

## Multiple linear regression

We first perform a regression model: `Age~Intercept+Weight+Height` in our function. Note that in this model, we include the intercept, and do not consider the interaction.

```{r}
lm1.2 <- lr(Age~Weight+Height, data)
```

And then check the accuracy of our function by existing functions in R.

```{r}
lm1.1 <- lm(Age~Weight+Height, data)

names(lm1.2)
# fitted values
all.equal(as.numeric(lm1.1$fitted.values), as.numeric(lm1.2$y.fitted))
# conventional and internally studendized residuals
all.equal(as.numeric(lm1.1$residuals), as.numeric(lm1.2[["residual"]][, 1]))
all.equal(as.numeric(rstandard(lm1.1)), as.numeric(lm1.2[["residual"]][, 2]))
# coefficients
all.equal(as.numeric(summary(lm1.1)$coefficients), as.numeric(lm1.2$cov.matrix))
# degree of freedom
all.equal(as.numeric(summary(lm1.1)$r.squared), as.numeric(lm1.2$r.squared))
# sigma
all.equal(as.numeric(summary(lm1.1)$sigma), as.numeric(lm1.2$sigma))
# F-test
all.equal(as.numeric(as.matrix(anova(lm(Age~1, data), lm1.1))[2, 5]), as.numeric(lm1.2$f.result[1]))
```

Additionally, we check the efficiency of our function.

```{r}
system.time({
  lm1.1 <- lm(Age~Weight+Height, data)
  anova(lm(Age~1, data), lm1.1)
  rstandard(lm1.1)
})
system.time(lm1.2 <- lr(Age~Weight+Height, data))
```

We can conclude that our function does not improve the efficiency of the original ones.

## Intercept

We then perform a regression model without intercept: `Age~Weight+Height` in our function. Note that in this model, we still exclude the interaction.

```{r}
lm2.2 <- lr(Age~Weight+Height, data, intercept = FALSE)
```

And then check the accuracy of our function by existing functions in R.

```{r}
lm2.1 <- lm(Age~-1+Weight+Height, data)

names(lm2.2)
# fitted values
all.equal(as.numeric(lm2.1$fitted.values), as.numeric(lm2.2$y.fitted))
# conventional and internally studendized residuals
all.equal(as.numeric(lm2.1$residuals), as.numeric(lm2.2[["residual"]][, 1]))
all.equal(as.numeric(rstandard(lm2.1)), as.numeric(lm2.2[["residual"]][, 2]))
# coefficients
all.equal(as.numeric(summary(lm2.1)$coefficients), as.numeric(lm2.2$cov.matrix))
# degree of freedom
all.equal(as.numeric(summary(lm2.1)$r.squared), as.numeric(lm2.2$r.squared))
# sigma
all.equal(as.numeric(summary(lm2.1)$sigma), as.numeric(lm2.2$sigma))
```

Additionally, we check the efficiency of our function.

```{r}
system.time({
  lm2.1 <- lm(Age~Weight+Height, data)
  anova(lm2.1)
  rstandard(lm2.1)
})
system.time(lm2.2 <- lr(Age~Weight+Height, data, intercept = FALSE))
```

We can conclude that our function does not improve the efficiency of the original ones.

## Interaction

We then perform a regression model with interaction: `Age~Intercept+Weight+Height+Weight*Height` in our function. Note that in this model, we still include the intercept.

```{r}
lm3.2 <- lr(Age~Weight+Height, data, interact = TRUE)
```

And then check the accuracy of our function by existing functions in R.

```{r}
lm3.1 <- lm(Age~Weight*Height, data)

names(lm3.2)
# fitted values
all.equal(as.numeric(lm3.1$fitted.values), as.numeric(lm3.2$y.fitted))
# conventional and internally studendized residuals
all.equal(as.numeric(lm3.1$residuals), as.numeric(lm3.2[["residual"]][, 1]))
all.equal(as.numeric(rstandard(lm3.1)), as.numeric(lm3.2[["residual"]][, 2]))
# coefficients
all.equal(as.numeric(summary(lm3.1)$coefficients), as.numeric(lm3.2$cov.matrix))
# degree of freedom
all.equal(as.numeric(summary(lm3.1)$r.squared), as.numeric(lm3.2$r.squared))
# sigma
all.equal(as.numeric(summary(lm3.1)$sigma), as.numeric(lm3.2$sigma))
# F-test
all.equal(as.numeric(as.matrix(anova(lm(Age~1, data), lm3.1))[2, 5]), as.numeric(lm3.2$f.result[1]))
```

Additionally, we check the efficiency of our function.

```{r}
system.time({
  lm3.1 <- lm(Age~Weight*Height, data)
  anova(lm(Age~1, data), lm3.1)
  rstandard(lm3.1)
})
system.time(lm3.2 <- lr(Age~Weight+Height, data, interact = TRUE))
```

We can conclude that our function does not improve the efficiency of the original ones.

## Reference cell coding

Regarding the category variable `Race1`, we first consider reference cell coding to examine its relationship with `Age`.

```{r}
lm4.2 <- lr(Age~Race1, data, category = 1)
```

And then check the accuracy of our function by existing functions in R.

```{r}
lm4.1 <- lm(Age~factor(Race1), data)

names(lm4.2)
# fitted values
all.equal(as.numeric(lm4.1$fitted.values), as.numeric(lm4.2$y.fitted))
# conventional and internally studendized residuals
all.equal(as.numeric(lm4.1$residuals), as.numeric(lm4.2[["residual"]][, 1]))
all.equal(as.numeric(rstandard(lm4.1)), as.numeric(lm4.2[["residual"]][, 2]))
# coefficients (Note that the order of coefficients is different)
all.equal(as.numeric(summary(lm4.1)$coefficients[c(1:3), ]), as.numeric(lm4.2$cov.matrix[c(1:3), ]))
all.equal(as.numeric(summary(lm4.1)$coefficients[4, ]), as.numeric(lm4.2$cov.matrix[5, ]))
all.equal(as.numeric(summary(lm4.1)$coefficients[5, ]), as.numeric(lm4.2$cov.matrix[4, ]))
# degree of freedom
all.equal(as.numeric(summary(lm4.1)$r.squared), as.numeric(lm4.2$r.squared))
# sigma
all.equal(as.numeric(summary(lm4.1)$sigma), as.numeric(lm4.2$sigma))
# F-test
all.equal(as.numeric(as.matrix(anova(lm(Age~1, data), lm4.1))[2, 5]), as.numeric(lm4.2$f.result[1]))
```

Additionally, we check the efficiency of our function.

```{r}
system.time({
  lm4.1 <- lm(Age~factor(Race1), data)
  anova(lm(Age~1, data), lm4.1)
  rstandard(lm4.1)
})
system.time(lm4.2 <- lr(Age~Race1, data, category = 1))
```

We can conclude that our function does not improve the efficiency of the original ones.

## Cell means coding

Regarding the category variable `Race1`, we then consider cell means coding to examine its relationship with `Age`.

```{r}
lm5.2 <- lr(Age~Race1, data, category = 2)
```

And then check the accuracy of our function by existing functions in R.

```{r}
lm5.1 <- lm(Age~-1+factor(Race1), data)

names(lm5.2)
# fitted values
all.equal(as.numeric(lm5.1$fitted.values), as.numeric(lm5.2$y.fitted))
# conventional and internally studendized residuals
all.equal(as.numeric(lm5.1$residuals), as.numeric(lm5.2[["residual"]][, 1]))
all.equal(as.numeric(rstandard(lm5.1)), as.numeric(lm5.2[["residual"]][, 2]))
# coefficients (Note that the order of coefficients is different)
all.equal(as.numeric(summary(lm5.1)$coefficients[c(1:3), ]), as.numeric(lm5.2$cov.matrix[c(1:3), ]))
all.equal(as.numeric(summary(lm5.1)$coefficients[4, ]), as.numeric(lm5.2$cov.matrix[5, ]))
all.equal(as.numeric(summary(lm5.1)$coefficients[5, ]), as.numeric(lm5.2$cov.matrix[4, ]))
# degree of freedom
all.equal(as.numeric(summary(lm5.1)$r.squared), as.numeric(lm5.2$r.squared))
# sigma
all.equal(as.numeric(summary(lm5.1)$sigma), as.numeric(lm5.2$sigma))
```

Additionally, we check the efficiency of our function.

```{r}
system.time({
  lm5.1 <- lm(Age~--1+factor(Race1), data)
  anova(lm5.1)
  rstandard(lm5.1)
})
system.time(lm5.2 <- lr(Age~Race1, data, category = 2))
```

We can conclude that our function does not improve the efficiency of the original ones.
